---
title: "Stability Increases with Dispersal Diversity in the Newfoundland and Labrador Shelves (figures only)"
authors:
  - name: Danielle M.B. McRae
    orcid: 0009-0004-5354-8843
    email: daniellembmacrae@gmail.com
    corresponding: true
    affiliations: 
      - ref: con
  - name: Valentin Lucet
    orcid: 0000-0003-0268-818X
    affiliations: 
      - ref: wlu
  - name: Eric J. Pedersen
    orcid: 0000-0003-1016-540X
    affiliations: 
      - ref: con
  - name: Jennifer M. Sunday
    orcid: 0000-0001-9372-040X
    affiliations: 
      - ref: mcg
affiliations:
  - id: con
    name: Concordia University
  - id: wlu 
    name: Wilfrid Laurier University
  - id: mcg 
    name: McGill University
title-block-published: "Last updated"  
date: now
date-format: long
format: html
execute:
  echo: true
  warning: false
  message: false
  comment: "#>"
  fig-path: "../figures/"
  fig-dpi: 600
abstract: |
  To understand and predict how regional diversity changes under directed environmental pressure, there needs to be more focus on monitoring broad-scale spatially connected communities (metacommunities). While the importance of dispersal behaviours in shaping community dynamics has long been recognized, much less is known about how variation among species in dispersal behaviours may affect community stability and complexity. Theoretical and experimental metacommunity studies have shown that functional diversity in dispersal types (i.e.  dispersal diversity) can act to desynchronize population fluctuations, and may also be important for stabilizing community dynamics at regional scales. Here we test the hypothesis that communities with greater dispersal diversity have greater stability in biomass over time. We quantified functional variation in dispersal-linked traits of groundfish species across space and time of the Newfoundland and Labrador (NL) shelves from 1995 to 2017. Local communities with higher dispersal diversity, measured using functional evenness, tended to be more stable over five-year intervals. This relationship held even after controlling for local taxonomic diversity as measured by effective species number. Our findings provide support for the hypothesis that dispersal diversity increases stability, suggesting this variable is a useful measure of metacommunity function and monitoring under environmental pressure.
keywords: |
  dispersal; stability; diversity
---

<!-- This is the format for text comments that will be ignored during renderings. Do not put R code in these comments because it will not be ignored. -->

<!-- With the following code you can access and display values from the yml header above. -->

Keywords: `r rmarkdown::metadata$keywords`

<!-- The actual document text starts here: -->

# Introduction

Here is a citation [@Marwick2017]

# Background

# Methods

# Results

<!-- Here's some example analysis code: -->

```{r}
#| label: get-data
#| eval: false
# Note the path that we need to use to access our data files when rendering this document
my_data <- read.csv(here::here('analysis/data/raw_data/my_csv_file.csv'))
```

```{r}
#| label: fig-demo-plot
#| fig-cap: "A plot of random numbers"
plot(rnorm(10))
```

@fig-demo-plot shows how we can have a caption and cross-reference for a plot. Note that figure label and cross-references must both be prefixed with `fig-`

```{r}
#| label: demo-inline-code
#| echo: false
x <- round(pi, 2)
```

Here is an example of inline code `r x` in the middle of a sentence.

# Discussion

# Conclusion

# Acknowledgements

<!-- The following line inserts a page break  -->

\newpage

# References

<!-- The following line ensures the references appear here for the MS Word or HTML output files, rather than right at the end of the document (this will not work for PDF files):  -->

::: {#refs}
:::

\newpage

### Colophon

This report was generated on `r Sys.time()` using the following computational environment and dependencies:

```{r}
#| label: colophon
#| cache: false

# which R packages and versions?
if ("devtools" %in% installed.packages()) devtools::session_info()
```

The current Git commit details are:

```{r}
# what commit is this file at? 
if ("git2r" %in% installed.packages() & git2r::in_repository(path = ".")) git2r::repository(here::here())  
```
